Structural Divergence Analysis

### 2017-04-09 analysis-dynamical-nH.Rmd
dinámica 9/4/17

In this section we consider dynamical divergence using normal modes coordinates.

```{r comparisson-of-means, fig.cap = "blue = experimental - green = mut - red = medium selection"}
# calculate means
mean.nH.exp = colMeans(nH.exp, na.rm = T)
mean.nH.mut = colMeans(nH.mut)
mean.nH.medium = colMeans(nH.medium)

@@

cor.df = data.frame("comparison" = c("exp vs mut", "exp vs medium", "mut vs medium"),
                    "R" = c(r.exp.mut, r.exp.medium, r.mut.medium))
knitr::kable(cor.df, digits = 2)
```

@@

plot(x = mean.nH.exp, 
     y = mean.nH.mut, 
     xlab = "exp.mean.nH", 
     ylab = "mut.mean.nH")
abline(0, 1)

plot(x = mean.nH.exp, 
     y = mean.nH.medium, 
     xlab = "exp.mean.nH", 
     ylab = "medium.mean.nH")
abline(0, 1)

plot(x = mean.nH.mut, 
     y = mean.nH.medium, 
     xlab = "mut.mean.nH", 
     ylab = "medium.mean.nH")
abline(0, 1)
```

```{r comparisson-of-means-100-modes}

@@

cor.df = data.frame("comparison" = c("exp vs mut", "exp vs medium", "mut vs medium"),
                    "R" = c(r.exp.mut, r.exp.medium, r.mut.medium))
knitr::kable(cor.df, digits = 2)

@@

plot(x = mean.nH.exp, 
     y = mean.nH.mut, 
     xlab = "exp.mean.nH", 
     ylab = "mut.mean.nH")
abline(0, 1)

plot(x = mean.nH.exp, 
     y = mean.nH.medium, 
     xlab = "exp.mean.nH", 
     ylab = "medium.mean.nH")
abline(0, 1)

plot(x = mean.nH.mut, 
     y = mean.nH.medium, 
     xlab = "mut.mean.nH", 
     ylab = "medium.mean.nH")
abline(0, 1)
```

>>>>

In this section we consider dynamical divergence using normal modes coordinates. We calculated mean nH and we z-normalized profiles.

```{r comparison-of-means, fig.cap = "blue = experimental - green = no-selection - red = selection"}
# calculate means
mean.nH.exp = colMeans(nH.exp, na.rm = T)
mean.nH.mut = colMeans(nH.mut)
mean.nH.medium = colMeans(nH.medium)

@@

cor.df = data.frame("comparison" = c("exp vs no-selection", "exp vs selection", "no-selection vs selection"),
                    "R" = c(r.exp.mut, r.exp.medium, r.mut.medium))
knitr::kable(cor.df, digits = 2)
```

@@

plot(x = mean.nH.exp, 
     y = mean.nH.mut, 
     xlab = "exp.mean.nH", 
     ylab = "no.selection.mean.nH")
abline(0, 1)

plot(x = mean.nH.exp, 
     y = mean.nH.medium, 
     xlab = "exp.mean.nH", 
     ylab = "selection.mean.nH")
abline(0, 1)

plot(x = mean.nH.mut, 
     y = mean.nH.medium, 
     xlab = "no.selection.mean.nH", 
     ylab = "selection.mean.nH")
abline(0, 1)
```

```{r comparison-of-means-100-modes}

@@

cor.df = data.frame("comparison" = c("exp vs no-selection", "exp vs selection", "no-selection vs selection"),
                    "R" = c(r.exp.mut, r.exp.medium, r.mut.medium))
knitr::kable(cor.df, digits = 2)

@@

plot(x = mean.nH.exp, 
     y = mean.nH.mut, 
     xlab = "exp.mean.nH", 
     ylab = "no.selection.mean.nH")
abline(0, 1)

plot(x = mean.nH.exp, 
     y = mean.nH.medium, 
     xlab = "exp.mean.nH", 
     ylab = "selection.mean.nH")
abline(0, 1)

plot(x = mean.nH.mut, 
     y = mean.nH.medium, 
     xlab = "no.selection.mean.nH", 
     ylab = "selection.mean.nH")
abline(0, 1)
```

### 2017-04-09 analysis-dynamical-nR.Rmd
dinámica 9/4/17

In this section we consider dynamical divergence using normal modes coordinates.

```{r comparisson-of-means, fig.cap = "blue = experimental - green = mut - red = medium selection"}
# calculate means
mean.nR.exp = colMeans(nR.exp, na.rm = T)
mean.nR.mut = colMeans(nR.mut)
mean.nR.medium = colMeans(nR.medium)

@@

plot(x = mean.nR.exp, 
     y = mean.nR.mut, 
     xlab = "exp.mean.nR", 
     ylab = "mut.mean.nR")
abline(0, 1)

plot(x = mean.nR.exp, 
     y = mean.nR.medium, 
     xlab = "exp.mean.nR", 
     ylab = "medium.mean.nR")
abline(0, 1)

plot(x = mean.nR.mut, 
     y = mean.nR.medium, 
     xlab = "mut.mean.nR", 
     ylab = "medium.mean.nR")
abline(0, 1)
```

```{r comparisson-of-means-100-modes}

@@

plot(x = mean.nR.exp, 
     y = mean.nR.mut, 
     xlab = "exp.mean.nR", 
     ylab = "mut.mean.nR")
abline(0, 1)

plot(x = mean.nR.exp, 
     y = mean.nR.medium, 
     xlab = "exp.mean.nR", 
     ylab = "medium.mean.nR")
abline(0, 1)

plot(x = mean.nR.mut, 
     y = mean.nR.medium, 
     xlab = "mut.mean.nR", 
     ylab = "medium.mean.nR")
abline(0, 1)
```

>>>>

In this section we consider dynamical divergence using normal modes coordinates. We calculated mean nR and we z-normalized profiles.

```{r comparison-of-means, fig.cap = "blue = experimental - green = no-selection - red = selection"}
# calculate means
mean.nR.exp = colMeans(nR.exp, na.rm = T)
mean.nR.mut = colMeans(nR.mut)
mean.nR.medium = colMeans(nR.medium)

@@

plot(x = mean.nR.exp, 
     y = mean.nR.mut, 
     xlab = "exp.mean.nR", 
     ylab = "no.selection.mean.nR")
abline(0, 1)

plot(x = mean.nR.exp, 
     y = mean.nR.medium, 
     xlab = "exp.mean.nR", 
     ylab = "selection.mean.nR")
abline(0, 1)

plot(x = mean.nR.mut, 
     y = mean.nR.medium, 
     xlab = "no.selection.mean.nR", 
     ylab = "selection.mean.nR")
abline(0, 1)
```

```{r comparison-of-means-100-modes}

@@

plot(x = mean.nR.exp, 
     y = mean.nR.mut, 
     xlab = "exp.mean.nR", 
     ylab = "no.selection.mean.nR")
abline(0, 1)

plot(x = mean.nR.exp, 
     y = mean.nR.medium, 
     xlab = "exp.mean.nR", 
     ylab = "selection.mean.nR")
abline(0, 1)

plot(x = mean.nR.mut, 
     y = mean.nR.medium, 
     xlab = "no.selection.mean.nR", 
     ylab = "selection.mean.nR")
abline(0, 1)
```

### 2017-04-09 analysis-structure.Rmd
dinámica 9/4/17

```{r set-my-colors}
# set my colors
myColors = c("blue", "darkgreen", "darkgreen", "red", "red") 
names(myColors) <- c("exp", "mut", "no-selection", "strong", "selection")
```

@@

## profiles to compare
dri2.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.norm.dr.squarei.csv", sep = ""))
dri2.strong.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_strong.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))
dri2.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))

## consurf
score.fname = file.path(data.dir, paste(p.ref, "_consurf.csv", sep = ""))
sim.data.no.sel.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_out_df.data.csv", sep = ""))
sim.data.strong.sel.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_strong.sel_out_df.data.csv", sep = ""))

@@

## strong
dri2.strong = read.csv(dri2.strong.fname, header = T)
site = site.info$site
dri2.strong = dri2.strong[1:length(protein.exp), site]
rownames(dri2.strong) = paste("strong", rownames(dri2.strong), sep = "")
```

@@

## strong
tdri2.strong = as.data.frame(t(as.matrix(dri2.strong)))
tdri2.strong = cbind(site.info, tdri2.strong)
siteComparison.strong = melt(tdri2.strong, id.vars = names(site.info), variable.name = "protein", value.name = "dri2")

# bluild the data.frame
siteComparison.all.datasets = rbind(data.frame("dataset" = "exp", siteComparison.exp),
                                    data.frame("dataset" = "mut", siteComparison.mut),
                                    data.frame("dataset" = "strong", siteComparison.strong))

@@

```{r-mean-se-z.RSD-vs-site, fig.cap = "Black lines are the active sites and grey lines are their contacts"}
# comparison of z.RSD profiles with functional sites
## prepare dat
dat = siteComparison.all.datasets
dat$dataset = revalue(dat$dataset, c("strong" = "selection"))
dat$dataset = revalue(dat$dataset, c("mut" = "no-selection"))

@@

# calculate correlations berween profiles
d.exp = subset(d, dataset == "exp")
d.mut = subset(d, dataset == "no-selection")
d.strong = subset(d, dataset == "selection")

# calculate cors
r.exp.mut = cor(d.exp$mean.z.RSD, d.mut$mean.z.RSD)
r.exp.strong = cor(d.exp$mean.z.RSD, d.strong$mean.z.RSD)
r.mut.strong = cor(d.mut$mean.z.RSD, d.strong$mean.z.RSD)

cor.df = data.frame("comparison" = c("exp vs mut", "exp vs strong", "mut vs strong"),
                    "R" = c(r.exp.mut, r.exp.strong, r.mut.strong))

@@

plot(x = d.exp$mean.z.RSD, 
     y = d.strong$mean.z.RSD, 
     xlab = "exp.mean.z.RSD", 
     ylab = "strong.mean.z.RSD", 
     xlim = c(min(c(d.exp$mean.z.RSD, d.strong$mean.z.RSD)), max(c(d.exp$mean.z.RSD, d.strong$mean.z.RSD))),
     ylim = c(min(c(d.exp$mean.z.RSD, d.strong$mean.z.RSD)), max(c(d.exp$mean.z.RSD, d.strong$mean.z.RSD))))
points(x = as.numeric(d.exp$mean.z.RSD)[contacts.active.sites], y = as.numeric(d.strong$mean.z.RSD)[contacts.active.sites], col = "green")
points(x = as.numeric(d.exp$mean.z.RSD)[active.sites], y = as.numeric(d.strong$mean.z.RSD)[active.sites], col = "yellow")
abline(0, 1)

plot(x = d.mut$mean.z.RSD, 
     y = d.strong$mean.z.RSD, 
     xlab = "mut.mean.z.RSD", 
     ylab = "strong.mean.z.RSD", 
     xlim = c(min(c(d.mut$mean.z.RSD, d.strong$mean.z.RSD)), max(c(d.mut$mean.z.RSD, d.strong$mean.z.RSD))),
     ylim = c(min(c(d.mut$mean.z.RSD, d.strong$mean.z.RSD)), max(c(d.mut$mean.z.RSD, d.strong$mean.z.RSD))))
points(x = as.numeric(d.mut$mean.z.RSD)[contacts.active.sites], y = as.numeric(d.strong$mean.z.RSD)[contacts.active.sites], col = "green")
points(x = as.numeric(d.mut$mean.z.RSD)[active.sites], y = as.numeric(d.strong$mean.z.RSD)[active.sites], col = "yellow")
abline(0, 1)
```

@@

## mut + sel / strong
strong.mean.z.RSD.vec <- vec2resno(as.vector(d.strong$mean.z.RSD), sort(rep(1:ncol(pdb.xyz.ca),3)))
write.pdb(xyz = c(pdb.xyz.ca), b = strong.mean.z.RSD.vec, file = file.path(data.dir, paste(p.ref, "_strong.pdb", sep = "")))
```

@@

## strong sel
sim.data.strong.sel = read.csv(sim.data.strong.sel.fname)
mut.site.strong.sel = sim.data.strong.sel$site
p.accept.strong.sel = sim.data.strong.sel$p.accept

# calculate mean.p.accept of simulated data
mean.p.accept.no.sel = c()
mean.p.accept.strong.sel = c()

# correct indexes
if (data.dir == "OUT/out_subset_CM_ANM") {
  mut.site.no.sel = mut.site.no.sel - n.sites
  mut.site.strong.sel = mut.site.strong.sel - n.sites
  CN.CM = sim.data.no.sel$CN
  CN = c()

@@

  p.accept.strong.sel.i = p.accept.strong.sel[which(mut.site.strong.sel == i)]
  mean.p.accept.strong.sel.i = mean(p.accept.strong.sel.i)
  mean.p.accept.strong.sel[i] = mean.p.accept.strong.sel.i
}

@@

# compare exp vs strong sel profiles
## plot
par(pty = "s")

plot(x = log(mean.p.accept.strong.sel), y = CN)
plot(x = log(mean.p.accept.strong.sel), y = score)
plot(x = score, y = CN)

## calculate corelations
r.mean.p.accept.strong.sel.CN = cor(log(mean.p.accept.strong.sel), CN)
r.mean.p.accept.strong.sel.score = cor(log(mean.p.accept.strong.sel), score)
r.score.CN = cor(score, CN)

cor.df = data.frame("comparison" = c("log(p.accept.strong.sel) vs CN", "log(p.accept.strong.sel) vs score", "score vs CN"),
                    "R" = c(r.mean.p.accept.strong.sel.CN , r.mean.p.accept.strong.sel.score, r.score.CN))

@@

plot(x = d.exp$mean.z.RSD, y = CN, xlab = "experimental <zRSDi>")
plot(x = d.mut$mean.z.RSD, y = CN, xlab = "theoretical with no selection <zRSDi>")
plot(x = d.strong$mean.z.RSD, y = CN, xlab = "theoretical with strong selection <zRSDi>")

## calculate corelations
r.exp.mean.z.RSD.CN = cor(d.exp$mean.z.RSD, CN)
r.mut.mean.z.RSD.CN = cor(d.mut$mean.z.RSD, CN)
r.strong.mean.z.RSD.CN = cor(d.strong$mean.z.RSD, CN)

cor.df = data.frame("comparison" = c("exp.mean.z.RSD vs CN", "mut.mean.z.RSD vs CN", "strong.mean.z.RSD vs CN"),
                    "R" = c(r.exp.mean.z.RSD.CN , r.mut.mean.z.RSD.CN, r.strong.mean.z.RSD.CN))

>>>>

```{r set-my-colors}
# set my colors
myColors = c("blue", "darkgreen", "darkgreen", "red", "red") 
names(myColors) <- c("exp", "mut", "no-selection", "medium", "selection")
```

@@

## profiles to compare
dri2.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.norm.dr.squarei.csv", sep = ""))
dri2.medium.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_medium.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))
dri2.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))

## consurf
score.fname = file.path(data.dir, paste(p.ref, "_consurf.csv", sep = ""))
sim.data.no.sel.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_out_df.data.csv", sep = ""))
sim.data.medium.sel.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_medium.sel_out_df.data.csv", sep = ""))

@@

## medium
dri2.medium = read.csv(dri2.medium.fname, header = T)
site = site.info$site
dri2.medium = dri2.medium[1:length(protein.exp), site]
rownames(dri2.medium) = paste("medium", rownames(dri2.medium), sep = "")
```

@@

## medium
tdri2.medium = as.data.frame(t(as.matrix(dri2.medium)))
tdri2.medium = cbind(site.info, tdri2.medium)
siteComparison.medium = melt(tdri2.medium, id.vars = names(site.info), variable.name = "protein", value.name = "dri2")

# bluild the data.frame
siteComparison.all.datasets = rbind(data.frame("dataset" = "exp", siteComparison.exp),
                                    data.frame("dataset" = "mut", siteComparison.mut),
                                    data.frame("dataset" = "medium", siteComparison.medium))

@@

```{r-mean-se-z.RSD-vs-site, fig.cap = "Black lines are the active sites and grey lines are their contacts"}
# comparison of z.RSD profiles with functional sites
## prepare dat
dat = siteComparison.all.datasets
dat$dataset = revalue(dat$dataset, c("medium" = "selection"))
dat$dataset = revalue(dat$dataset, c("mut" = "no-selection"))

@@

# calculate correlations berween profiles
d.exp = subset(d, dataset == "exp")
d.mut = subset(d, dataset == "no-selection")
d.medium = subset(d, dataset == "selection")

# calculate cors
r.exp.mut = cor(d.exp$mean.z.RSD, d.mut$mean.z.RSD)
r.exp.medium = cor(d.exp$mean.z.RSD, d.medium$mean.z.RSD)
r.mut.medium = cor(d.mut$mean.z.RSD, d.medium$mean.z.RSD)

cor.df = data.frame("comparison" = c("exp vs mut", "exp vs medium", "mut vs medium"),
                    "R" = c(r.exp.mut, r.exp.medium, r.mut.medium))

@@

plot(x = d.exp$mean.z.RSD, 
     y = d.medium$mean.z.RSD, 
     xlab = "exp.mean.z.RSD", 
     ylab = "medium.mean.z.RSD", 
     xlim = c(min(c(d.exp$mean.z.RSD, d.medium$mean.z.RSD)), max(c(d.exp$mean.z.RSD, d.medium$mean.z.RSD))),
     ylim = c(min(c(d.exp$mean.z.RSD, d.medium$mean.z.RSD)), max(c(d.exp$mean.z.RSD, d.medium$mean.z.RSD))))
points(x = as.numeric(d.exp$mean.z.RSD)[contacts.active.sites], y = as.numeric(d.medium$mean.z.RSD)[contacts.active.sites], col = "green")
points(x = as.numeric(d.exp$mean.z.RSD)[active.sites], y = as.numeric(d.medium$mean.z.RSD)[active.sites], col = "yellow")
abline(0, 1)

plot(x = d.mut$mean.z.RSD, 
     y = d.medium$mean.z.RSD, 
     xlab = "mut.mean.z.RSD", 
     ylab = "medium.mean.z.RSD", 
     xlim = c(min(c(d.mut$mean.z.RSD, d.medium$mean.z.RSD)), max(c(d.mut$mean.z.RSD, d.medium$mean.z.RSD))),
     ylim = c(min(c(d.mut$mean.z.RSD, d.medium$mean.z.RSD)), max(c(d.mut$mean.z.RSD, d.medium$mean.z.RSD))))
points(x = as.numeric(d.mut$mean.z.RSD)[contacts.active.sites], y = as.numeric(d.medium$mean.z.RSD)[contacts.active.sites], col = "green")
points(x = as.numeric(d.mut$mean.z.RSD)[active.sites], y = as.numeric(d.medium$mean.z.RSD)[active.sites], col = "yellow")
abline(0, 1)
```

@@

## mut + sel / medium
medium.mean.z.RSD.vec <- vec2resno(as.vector(d.medium$mean.z.RSD), sort(rep(1:ncol(pdb.xyz.ca),3)))
write.pdb(xyz = c(pdb.xyz.ca), b = medium.mean.z.RSD.vec, file = file.path(data.dir, paste(p.ref, "_medium.pdb", sep = "")))
```

@@

## medium sel
sim.data.medium.sel = read.csv(sim.data.medium.sel.fname)
mut.site.medium.sel = sim.data.medium.sel$site
p.accept.medium.sel = sim.data.medium.sel$p.accept

# calculate mean.p.accept of simulated data
mean.p.accept.no.sel = c()
mean.p.accept.medium.sel = c()

# correct indexes
if (data.dir == "OUT/out_subset_CM_ANM") {
  mut.site.no.sel = mut.site.no.sel - n.sites
  mut.site.medium.sel = mut.site.medium.sel - n.sites
  CN.CM = sim.data.no.sel$CN
  CN = c()

@@

  p.accept.medium.sel.i = p.accept.medium.sel[which(mut.site.medium.sel == i)]
  mean.p.accept.medium.sel.i = mean(p.accept.medium.sel.i)
  mean.p.accept.medium.sel[i] = mean.p.accept.medium.sel.i
}

@@

# compare exp vs medium sel profiles
## plot
par(pty = "s")

plot(x = log(mean.p.accept.medium.sel), y = CN)
plot(x = log(mean.p.accept.medium.sel), y = score)
plot(x = score, y = CN)

## calculate corelations
r.mean.p.accept.medium.sel.CN = cor(log(mean.p.accept.medium.sel), CN)
r.mean.p.accept.medium.sel.score = cor(log(mean.p.accept.medium.sel), score)
r.score.CN = cor(score, CN)

cor.df = data.frame("comparison" = c("log(p.accept.medium.sel) vs CN", "log(p.accept.medium.sel) vs score", "score vs CN"),
                    "R" = c(r.mean.p.accept.medium.sel.CN , r.mean.p.accept.medium.sel.score, r.score.CN))

@@

plot(x = d.exp$mean.z.RSD, y = CN, xlab = "experimental <zRSDi>")
plot(x = d.mut$mean.z.RSD, y = CN, xlab = "theoretical with no selection <zRSDi>")
plot(x = d.medium$mean.z.RSD, y = CN, xlab = "theoretical with medium selection <zRSDi>")

## calculate corelations
r.exp.mean.z.RSD.CN = cor(d.exp$mean.z.RSD, CN)
r.mut.mean.z.RSD.CN = cor(d.mut$mean.z.RSD, CN)
r.medium.mean.z.RSD.CN = cor(d.medium$mean.z.RSD, CN)

cor.df = data.frame("comparison" = c("exp.mean.z.RSD vs CN", "mut.mean.z.RSD vs CN", "medium.mean.z.RSD vs CN"),
                    "R" = c(r.exp.mean.z.RSD.CN , r.mut.mean.z.RSD.CN, r.medium.mean.z.RSD.CN))
